Multi-Source Debiased Calibration Simulation with Estimated Alpha

### 2024-11-04 src/explore/20241101-msdcsim_estalpha.qmd
Added real data simulation. Computation still needs work. It takes too long to transfer data.

---
title: "Multi-Source Debiased Calibration Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

@@

This report simulates multi-source sampling with the debiased
calibration of @kwon2024debiased. We check for the following:

1. Bias of $\hat Y$ and,
2. Bias of $\hat V(\hat Y)$.

@@

    if (entropy == "EL") {
      w <- -d2i / drop(zmat %*% lam)
    } else if (entropy == "ET") {
      w <- d2i * exp(drop(zmat %*% lam))
    } else {
      stop("We only accept entropy of EL or ET.")
    }

    if (returnw) {
      return(w)
    } else {
      return(drop((w * w1i * qi) %*% zmat) - T1)
    }
  }

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    if (entropy == "EL") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

@@

  init <- c(rep(0, ncol(zmat) - 1), 1)
  res <- 
  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
          T1 = T1, qi = qi, entropy = entropy,
          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))

@@

dc_ms_ybar <- 
  function(p0_df, p1_df, p2_df, pop_df, estT1, qi = 1, entropy = "EL", reg = FALSE) {

  d_vec <- 1 / (p0_df$pi0)
  if (entropy == "EL") {
    gdi <- -1 / d_vec
    gdip <- -1 / (1 / pop_df$pi0)
    gpinv <- d_vec^2
  } else if (entropy == "ET") {
    gdi <- log(d_vec)
    gdip <- log(1 / pop_df$pi0)
    gpinv <- d_vec
  } else {
    stop("This type of entropy has not yet been implemented.")
  }

  samp_X <- cbind(rep(1, nrow(p0_df)), as.matrix(select(p0_df, X1, X2, X3)))
  zmat <- cbind(samp_X, gdi)
  w1i <- 1
  d0i <- 1 / p0_df$pi0

  if (estT1) {
    x1_0 <- sum(p0_df$X1 / p0_df$pi0) / N_obs
    x2_0 <- sum(p0_df$X2 / p0_df$pi0) / N_obs
    x3_0 <- sum(p0_df$X3 / p0_df$pi0) / N_obs

@@

    T1 <- c(nrow(pop_df),
            xgls[1] * nrow(pop_df),
            xgls[2] * nrow(pop_df),
            xgls[3] * nrow(pop_df),
            sum(gdip * qi))
  } else {
    T1 <- c(nrow(pop_df),
            sum(pop_df$X1),
            sum(pop_df$X2),
            sum(pop_df$X3),
            sum(gdip * qi))
  }

  dc_w <- solveGE(zmat, w1i, d0i, T1, qi, entropy)

  # Mean Estimation
  theta <- sum(p0_df$Y * dc_w * w1i) / nrow(pop_df)
  z_dc <- model.matrix(~1 + p0_df$X1 + p0_df$X2 + p0_df$X3 + gdi)
  gam_dc <- 
    solve(t(z_dc) %*% diag(d0i * gpinv * qi) %*% z_dc,
          t(z_dc) %*% diag(d0i * gpinv) %*% p0_df$Y)
  coefs <- as.numeric(gam_dc)
  u_mat <- 
  matrix(c(rep(1, nrow(p0_df)), p0_df$X1, p0_df$X2, p0_df$X3, gdi * qi), ncol = 5)
  eps <- (p0_df$Y - drop(u_mat %*% coefs))
  if (estT1) {
    # Variance Estimation
    # This is the variance of bar x gls not x gls total.
    var_x <- MASS::ginv(t(x_mat) %*% MASS::ginv(vmat) %*% x_mat)

    v1 <- matrix(coefs[2:4], nrow = 1) %*% var_x %*% matrix(coefs[2:4], ncol = 1)
    v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / nrow(pop_df)^2
    v_hat <- v1 + v2 
  } else {
    v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / nrow(pop_df)^2
    v_hat <- v2
  }
  if (reg) {
    theta <- 
    (drop(T1 %*% gam_dc) + 
      sum(1 / p0_df$pi0 * (p0_df$Y - drop(z_dc %*% gam_dc)))) / nrow(pop_df) 
  }

@@

    ht <- ht_est(p0_df, N_obs)
    reg <- reg_est(p0_df, p1_df, N_obs)
    ms_pop <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = FALSE)
    ms_est <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = TRUE)
    ms_reg <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = TRUE, reg = TRUE)

    return(tibble(Est = c("HT", "NNReg", "MSPop", "MSEst", "MSReg"),
                  Theta = c(ht[[1]], reg[[1]], ms_pop[[1]], ms_est[[1]], ms_reg[[1]]),
                  Var = c(ht[[2]], reg[[2]], ms_pop[[2]], ms_est[[2]], ms_reg[[2]]),
                  Iter = iter)
    )

@@

# Check for failures
filter(mc_res, is.na(Theta))

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest) %>%
  mutate(Ind = case_when(
    Est == "MSReg" ~ 5,
    Est == "MSEst" ~ 4,
    Est == "MSPop" ~ 3,
    Est == "NNReg" ~ 2,
    Est == "HT" ~ 1)
  ) %>%
  arrange(Ind) %>%
  select(Est, Bias, RMSE, EmpCI, Ttest)

%>%
  knitr::kable("latex", booktabs = TRUE, digits = 4) %>%
  cat(file = "tables/msdcsim1.tex")

>>>>

---
title: "Multi-Source Debiased Calibration Simulation with Estimated Alpha"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

@@

This report simulates multi-source sampling with the debiased
calibration of @kwon2024debiased with an estimated $\alpha$ value. 
With this modification we do not need the design weights for the population.

@@

    if (entropy == "EL") {
      w <- (-d2i) / drop(zmat %*% lam)
    } else if (entropy == "ET") {
      w <- d2i * exp(drop(zmat %*% lam))
    } else {
      stop("We only accept entropy of EL or ET.")
    }

    if (returnw) {
      return(w)
    } else {
      return(T1 - drop((w * w1i * qi) %*% zmat))
    }
  }

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    end <- ncol(zmat)
    if (entropy == "EL") {
      res <- (-1) * t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- (-1) * t(zmat) %*% 
        diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

@@

  # zmat still contains g(di) as the last column
  init <- c(rep(0, ncol(zmat) - 1), 1)
  res <- 
  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
          T1 = T1, qi = qi, entropy = entropy,
          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))

@@

# The solveGE_alpha function estimates alpha and the weights.
# This is useful for when we do not have design weights for the entire
# population.
solveGE_alpha <- function(alpha, zmat, w1i, d2i, T1, qi, entropy, retw = FALSE) {

  N <- T1[1]
  T1_adj <- T1
  T1_adj[length(T1_adj)] <- alpha * N
  dc_w <- solveGE(zmat, w1i, d2i, T1_adj, qi, entropy)

  if (retw) {
    return(dc_w)
  }

  Ka_type <- "alpha"
  if (Ka_type == "alpha") {
    # HACK: G(x) = -log(x) for EL only
    G <- function(x) {-log(x)}
    return(sum(G(dc_w)) - N * alpha)
  }
}

dc_ms_ybar <- 
  function(p0_df, p1_df, p2_df, pop_df, qi = 1, entropy = "EL", reg = FALSE) {

  # Since we have a SRS, it is fine to use this info.
  N_obs <- nrow(pop_df)
  d_vec <- 1 / (p0_df$pi0)
  if (entropy == "EL") {
    gdi <- -1 / d_vec
    gpinv <- d_vec^2
  } else if (entropy == "ET") {
    gdi <- log(d_vec)
    gdip <- log(1 / pop_df$pi0)
    gpinv <- d_vec
  } else {
    stop("This type of entropy has not yet been implemented.")
  }

  samp_X <- model.matrix(~X1 + X2 + X3, data = p0_df)
  zmat <- cbind(samp_X, gdi)
  w1i <- 1

  x1_0 <- sum(p0_df$X1 / p0_df$pi0) / N_obs
  x2_0 <- sum(p0_df$X2 / p0_df$pi0) / N_obs
  x3_0 <- sum(p0_df$X3 / p0_df$pi0) / N_obs

@@

  T1 <- c(N_obs,
          xgls[1] * N_obs,
          xgls[2] * N_obs,
          xgls[3] * N_obs,
          NA)

  # Estimating Alpha
  alpha_init <- (-1) * nrow(p0_df) / N_obs
  a_res <- 
    nlm(f = solveGE_alpha, p = alpha_init,
        zmat = zmat, w1i = w1i, d2i = d_vec, T1 = T1, qi = qi, entropy = entropy)

  if (a_res$code != 1) {
    warning(paste0("a_res has code", a_res$code))
  }

  dc_w <- 
    solveGE_alpha(a_res$estimate, zmat, w1i, d_vec, T1, qi, entropy, retw = TRUE)

  # Mean Estimation
  # Do we want N_obs? Or an estimated N_hat? 
  # N_obs is probably fine since we have SRS.
  theta <- sum(p0_df$Y * dc_w * w1i) / N_obs
  z_dc <- samp_X
  gam_dc <- 
    solve(t(z_dc) %*% diag(d_vec * gpinv * qi) %*% z_dc,
          t(z_dc) %*% diag(d_vec * gpinv) %*% p0_df$Y)
  coefs <- as.numeric(gam_dc)
  u_mat <- 
  matrix(c(rep(1, nrow(p0_df)), p0_df$X1, p0_df$X2, p0_df$X3), ncol = 4)
  eps <- (p0_df$Y - drop(u_mat %*% coefs))
  # Variance Estimation
  # This is the variance of bar x gls not x gls total.
  var_x <- MASS::ginv(t(x_mat) %*% MASS::ginv(vmat) %*% x_mat)

  v1 <- matrix(coefs[2:4], nrow = 1) %*% var_x %*% matrix(coefs[2:4], ncol = 1)
  v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / nrow(pop_df)^2
  v_hat <- v1 + v2 
  if (reg) {
    theta <- 
    (drop(T1 %*% gam_dc) + 
      sum(1 / p0_df$pi0 * (p0_df$Y - drop(z_dc %*% gam_dc)))) / nrow(pop_df) 
  }

@@

    ht <- ht_est(p0_df, N_obs)
    reg <- reg_est(p0_df, p1_df, N_obs)
    ms_est <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df)

    return(tibble(Est = c("HT", "NNReg", "MSEst"),
                  Theta = c(ht[[1]], reg[[1]], ms_est[[1]]), 
                  Var = c(ht[[2]], reg[[2]], ms_est[[2]]), 
                  Iter = iter)
    )

@@

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            SE = sd(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE),
            RelBias = (EstVar - MCVar) / MCVar) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest) %>%
  mutate(Ind = case_when(
    Est == "MSReg" ~ 5,
    Est == "MSEst" ~ 4,
    Est == "MSPop" ~ 3,
    Est == "NNReg" ~ 2,
    Est == "HT" ~ 1)
  )

%>%
  knitr::kable("latex", booktabs = TRUE, digits = 4) %>%
  cat(file = "tables/msdc_estalphasim1.tex")
